---
title: "NFL Reports"
author: "Justin Post"
output: html_document  
params: 
      team: "Pittsburgh Steelers"
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

options(digits = 2)
library(tidyverse)
library(DT)
```


#Read in Team Data

First we need to load the team data.  In the YAML header we created a `team` parameter.  We will load the full data and then use the `params$team` to access it.  These parameters **cannot** be modified using R code!  

## Data read in 

We'll use the `readr` and `dplyr` packags from the `tidyverse`.  The `filter` here just selects data from the `r params$team`.  

```{r getData}
NFLData <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/scoresFull.csv")

params$team

teamData <- filter(NFLData, (awayTeam == params$team) | (homeTeam == params$team))
```

Selected columns from the data set are given below.  

```{r printData}
datatable(select(teamData, week, season, awayTeam, AQ2, AFinal, HQ2, HFinal))
```


#Obtain Yearly Summaries

We'll look at just the difference in offensive statistics and defensive statistics (for instance, total yards - total yards given up) for each game.  This will be broken up by year and by home/away status.  


```{r offensiveSummaries}
homeStats <- teamData %>% filter(homeTeam == params$team) %>% transmute(season = season, Q2Diff = HQ2 - AQ2, scoreDiff = HFinal - AFinal, yardDiff = HtotalYds - AtotalYds, turnoverDiff = Hturnovers - Aturnovers, playDiff = HtotalPlays - AtotalPlays)

awayStats <- teamData %>% filter(awayTeam == params$team) %>% transmute(season = season, Q2Diff = AQ2 - HQ2, scoreDiff = AFinal - HFinal, yardDiff = AtotalYds - HtotalYds, turnoverDiff = Aturnovers - Hturnovers, playDiff = AtotalPlays - HtotalPlays)

homeStats$HomeAway <- rep("Home", length(homeStats$season))

awayStats$HomeAway <- rep("Away", length(awayStats$season))

stats <- rbind(homeStats, awayStats)
```

Now find the averages.  

```{r create-avgData}
avg <- stats %>% 
	group_by(season, HomeAway) %>% 
	summarise(Q2Avg = mean(Q2Diff), finalAvg = mean(scoreDiff), yardAvg = mean(yardDiff), turnoverAvg = mean(turnoverDiff), playAvg = mean(playDiff)) %>% 
	ungroup()

avg[, -1:-2] <- round(select(avg, Q2Avg:playAvg), 1)
```

The data we'll have to work with for plotting is given below.  

```{r avgData}
datatable(avg)
```

# Plots over season  

Now we can create plots to see how the teams have done over the years.  

```{r, echo = FALSE}
#stack data to make legend easier
scorePlotData <- data.frame(season = rep(avg$season, 2), 
														scoreDiff = c(avg$Q2Avg, avg$finalAvg), 
														gameTime = c(rep("2nd Q", length(avg$Q2Avg)),
																				 rep("Final", length(avg$Q2Avg))), 
														HomeAway = rep(avg$HomeAway, 2))

scoresPlot <- ggplot(data = scorePlotData, aes(x = season, y = scoreDiff, color = HomeAway)) +
	geom_line(aes(linetype =  gameTime), lwd = 1) +
	ggtitle(label = paste0(params$team, " Score Differences"))

scoresPlot
```
