---
title: "Analysis of River Data Set"
author: "your_name_here"
date: "August 2018"
output:
  html_document:  
    toc: true  
    code_folding: hide  
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(knitr)
```

# About the Data We'll Use  

This presentation will represent a replication of some analysis done in the article [*Control factors and scale analysis of annual river water, sediments and carbon transport in China*](http://www.nature.com/articles/srep25963).  The data are available to download there but we will link to in on the gitHub web site for this course.  A description of the variables is available at <http://www.nature.com/articles/srep25963/tables/1>.  

<br>
<br>
<br>

#About the Analysis We'll Look At  

## Numerical Summaries  

The analysis will go through things like finding **means**, **counts**, and **standard deviations**.  These summaries will be found for the variables below:  

 - Rc (Runoff coefficient - runoff depth divided by annual rainfall RD/total rainfall)  
 - TSSC_mgPl (Total suspended sediment concentration (mg/L)	- average total suspended sediment concentration in the runoff)  
 
Means will be found at **each** level of a new *size* variable.  The variable is created using the `Size_km2` (watershed, catchment or basin area in km^2^) variable as described in the table below.  

Size in km^2^ | Size Classification 
--------------|--------------------
< 15000       | 'Small'
< 100,000     | 'Medium'
< 350,000     | 'Sizeable'
< 700,000     | 'Large'
>= 700,000    | 'Great'
 
## Graphical Summaries  
 
Graphs will be made to visualize the data as well.  
 
 1. Boxplots  
 2. Histograms
 3. Scatter Plots  

<br>
<br>
<br>

# Reading in Data

We'll load the `readr` package to read in the data.

```{r}
library(readr)
riverData <- read_csv("https://raw.githubusercontent.com/jbpost2/IntermediateR/master/datasets/river.csv")
```

To create the new size variable we'll use the `ifelse` command and `dplyr` functions.  Then we'll order the levels for plotting purposes.

```{r}
library(dplyr)
riverData <- riverData %>% 
	mutate(Size = ifelse(Size_km2 < 15000, "Small", 
											 ifelse(Size_km2 < 100000, "Medium", 
											 			 ifelse(Size_km2 < 350000, "Sizeable", 
											 			 			 ifelse(Size_km2 < 700000, "Large", "Great")))))

riverData$Size <- ordered(riverData$Size, levels = c("Small", "Medium", "Sizeable", "Large", "Great"))
```

Later we'll use a binned version of the MAT variable (Mean annual temperature (degrees C)	- long-term mean annual temperature minus average temperature over 30 years or more) created here and also reordered.

```{r}
riverData <- riverData %>% 
	mutate(MAT_Class = ifelse(MAP < 600, "Semiarid",
														ifelse(MAP >= 600 & MAP < 850, "Moist", 
																	 ifelse(MAP >= 850 & MAP < 1500, "Humid", "Wet"))))

riverData$MAT_Class <- ordered(riverData$MAT_Class, levels = c("Semiarid", "Moist", "Humid", "Wet"))
```

<br>
<br>
<br>

#Analysis Part I  

## The Rc Variable

The `dplyr` package will be used for data manipulation and basic numeric summaries.  The means, counts, and standard deviations for each level of Size are given below.  

```{r}
riverData %>% group_by(Size) %>% summarise(mean(Rc, na.rm = TRUE)) %>% kable()
riverData %>% group_by(Size) %>% summarise(count = sum(!is.na(Rc))) %>% kable()
riverData %>% group_by(Size) %>% summarise(sd = sd(Rc, na.rm = TRUE)) %>% kable()
```

## The TSSC Variable

The means, counts, and standard deviations for each level of Size are given below.  

```{r}
riverData %>% group_by(Size) %>% summarise(mean(TSSC_mgPl, na.rm = TRUE)) %>% kable()
riverData %>% group_by(Size) %>% summarise(count = sum(!is.na(TSSC_mgPl))) %>% kable()
riverData %>% group_by(Size) %>% summarise(sd = sd(TSSC_mgPl, na.rm = TRUE)) %>% kable()
```

<br>
<br>
<br>

# Analysis Part II  

## Boxplots

The `ggplot2` package will be used to create some fun plots to visualize the data.  First, a boxplot of the Rc variable for each level of Size where the means of each group are connected.  

```{r, warning = TRUE, message = TRUE}
library(ggplot2)
g <- ggplot(data = riverData, aes(x = Size, y = Rc))
g + geom_boxplot(lwd = 1) + 
	coord_cartesian(ylim = c(0, 1)) + 
	labs(title = "Varying Trends of Rc Across Sizes", x = "Size") + 
	geom_point(aes(col = MAT_Class), size = 4) + 
	stat_summary(fun.y = mean, geom = "line", lwd = 2, aes(group = MAT_Class, col = MAT_Class))
```	
	
## Histograms  

Histograms of the Rc variable for each setting of Size are given below.

```{r, warning = TRUE, message = TRUE}
g <- ggplot(data = riverData, aes(x = Rc))
g + geom_histogram() + 
	facet_grid(. ~ Size)
```

## Scatter Plots

Lastly, scatter plots of the Rc variable and the TSSC variable are made *for each* combination of the Size and MAT_Class variables.  

```{r, warning = TRUE, message = TRUE}
g <- ggplot(data = riverData, aes(x = Rc, y = TSSC_mgPl))  
g + geom_point() + 
	facet_grid(Size ~ MAT_Class)
```

A 3D scatter plot.

```{r}
library(threejs)

scatterplot3js(x = riverData$Rc, y = riverData$TSSC_mgPl, z = riverData$MAT)
```

