---
title: "Grabbing and Cleaning NHL Data"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(RCurl)
library(rjson)
library(knitr)
opts_chunk$set(message = FALSE, cache = FALSE)
```

First get the game IDs from the NHL API.  Start with a single day to check out structure.

```{r}
oneDay <- fromJSON(getURL("https://statsapi.web.nhl.com/api/v1/schedule?startDate=2018-10-14&endDate=2018-10-14"))
str(oneDay)
```

Just want to grab the game IDs that they use so we can pull appropriate play by play and shift data.  Might as well grab some other things too I suppose!

Data of interest stored in first list element of `dates`:
```{r}
oneDay$dates[[1]]

```

Write a function to get what we want.
```{r}
extractNHLData <- function(gameStuff){
  info <- list()
  
  info$ID <- gameStuff$`gamePk`
  info$gameType <- gameStuff$`gameType`
  info$gameDayTime <- gameStuff$`gameDate`
  info$awayTeam <- gameStuff$`teams`$`away`$`team`$`name`
  info$awayTeamID <- gameStuff$`teams`$`away`$`team`$`id`
  info$awayScore <- gameStuff$`teams`$`away`$`score`
  info$awayWins <- gameStuff$`teams`$`away`$`leagueRecord`$`wins`
  info$awayLosses <- gameStuff$`teams`$`away`$`leagueRecord`$`losses`
  info$awayOT <- gameStuff$`teams`$`away`$`leagueRecord`$`ot`
  info$homeTeam <- gameStuff$`teams`$`home`$`team`$`name`
  info$homeTeamID <- gameStuff$`teams`$`home`$`team`$`id`
  info$homeScore <- gameStuff$`teams`$`home`$`score`
  info$homeWins <- gameStuff$`teams`$`home`$`leagueRecord`$`wins`
  info$homeLosses <- gameStuff$`teams`$`home`$`leagueRecord`$`losses`
  info$homeOT <- gameStuff$`teams`$`home`$`leagueRecord`$`ot`
  info$venue <- gameStuff$`venue`$`name`
  return(gameInfo = info)
  }

parsed <- t(sapply(FUN = extractNHLData, oneDay$dates[[1]]$`games`) )
parsed

```

Looking good.  This works for a particular day (`dates[[1]]`).  Need to make it work for everyday grabbed.  

```{r}
twoDays <- fromJSON(getURL("https://statsapi.web.nhl.com/api/v1/schedule?startDate=2018-10-14&endDate=2018-10-15"))
length(twoDays$dates)

parseSingleDay <- function(day, object){
  t(sapply(FUN = extractNHLData, object$dates[[day]]$`games`) )
}

parseSingleDay(2, twoDays)

temp <- lapply(X = 1:2, FUN = parseSingleDay, object = twoDays)
temp

#convert to data frames in each element and combine
lapply(temp, as.data.frame) %>% bind_rows(.id = "column_label")
```

Turn it all into one easy function, grab more dates, and test.

```{r}
combineDays <- function(object){
  lapply(X = 1:length(object$dates), FUN = parseSingleDay, object) %>% lapply(as.data.frame) %>% bind_rows(.id = "column_label")
}

combineDays(twoDays)

#full function
parseNHLData <- function(startDate = "2018-10-14", endDate = "2018-10-15"){
  url <- paste0("https://statsapi.web.nhl.com/api/v1/schedule?startDate=", startDate, "&endDate=", endDate)
  raw <- fromJSON(getURL(url))
  combineDays(raw)
}

NHLData <- parseNHLData(startDate = "2018-10-14", endDate = "2018-11-01")

NHLData
```

```{r, eval = FALSE}
#first day of season has some weird game between Oilers and KLH...
#just do as a special case
temp <- fromJSON(getURL("https://statsapi.web.nhl.com/api/v1/schedule?startDate=2018-10-03&endDate=2018-10-03"))
temp$dates[[1]]$`games`[[1]] <- NULL
NHLDay1 <- t(sapply(FUN = extractNHLData, temp$dates[[1]]$`games`))
NHLData <- parseNHLData(startDate = "2018-10-04", endDate = "2019-04-06")
NHLData <- NHLData %>% select(-V1:-V3, - column_label) 
NHLData <- rbind(NHLDay1, NHLData)
#can't write to file due to weird list columns...
dput(NHLData, file = "NHLGameData.txt")
```


Ok, now that the meta data is grabbed - ready to get the play-by-play and the shift data!

Check shift data
```{r, eval = FALSE}
temp2 <- fromJSON(getURL(paste0("http://www.nhl.com/stats/rest/shiftcharts?cayenneExp=gameId=",gameIDs[1])))
length(temp2$data)
t2 <- lapply(temp2$data, unlist)
t3 <- t2 %>% lapply(as.data.frame) %>% lapply(t)
t3 %>% bind_rows()
rbind(t3[[1]], t3[[833]])
#some are of differing lengths - appears to be when goals or something happens. Those are at the end though
```

Looks like detailCode of 0 means a shift I think.  Also may need to get player info later.
```{r, eval = FALSE}
temp2 <- fromJSON(getURL(paste0("http://www.nhl.com/stats/rest/shiftcharts?cayenneExp=gameId=",gameIDs[1]))) 
temp3 <- lapply(temp2$data, unlist)  
temp4 <- lapply(temp3, function(x){if(x["detailCode"] != "0"){return(NULL)} else{return(x)}})
temp5 <- lapply(temp4, as.data.frame) %>% lapply(t) %>% unlist() %>% matrix(ncol = 16, byrow = TRUE)

```

Ok that seems to get everything needed in the right form.  Write functions to get it for a given game.

```{r}
extractShiftData <- function(gameStuff){
  unlisted <- lapply(gameStuff, unlist)
  columns <- names(unlisted[[1]])
  formatted <- unlisted %>% lapply(function(x) { 
        if(x["detailCode"] != "0"){
          return(NULL)
        } else {
          return(x)
        }
      }) %>% 
    lapply(as.data.frame) %>% 
    lapply(t)
  #hopefully everything has length 16 (other than Null)
  #if not make it 16 to be safe
  checked <- formatted %>% lapply(function(x){
    if(length(x) != 16){
      x <- c(x, rep(NA, 16-length(x)))
    } else {
      x
    }
  })  %>% 
    unlist() %>% 
    matrix(ncol = 16, byrow = TRUE)
    colnames(checked) <- columns
    return(checked)
}

shiftGame <- fromJSON(getURL(paste0("http://www.nhl.com/stats/rest/shiftcharts?cayenneExp=gameId=",gameIDs[1])))

extractShiftData(shiftGame$data)
```

Now combining games.

```{r}
shiftGame2 <- fromJSON(getURL(paste0("http://www.nhl.com/stats/rest/shiftcharts?cayenneExp=gameId=",gameIDs[2])))

shift1 <- extractShiftData(shiftGame$data)
shift2 <- extractShiftData(shiftGame2$data)

#same structures, just bind_rows for all games found
gameIDs <- NHLData$ID

combineDays <- function(gameIDs){ 
  rawData <- lapply(gameIDs, 
         function(x){
           fromJSON(getURL(paste0("http://www.nhl.com/stats/rest/shiftcharts?cayenneExp=gameId=",x)))
         }
  )
  shifts <- lapply(rawData, function(x) { 
    extractShiftData(x$data)
    })
  shifts %>% lapply(as.data.frame, stringsAsFactors = FALSE) %>% bind_rows()
  }

shiftData <- combineDays(gameIDs[1:20])
write_csv(x = shiftData, path = "NHLShiftData.csv")
```



Let's parse some play by play.
```{r, eval = TRUE}
gameIDs <- unlist(NHLData$ID)
gameIDs[1]
temp <- fromJSON(getURL(paste0("https://statsapi.web.nhl.com/api/v1/game/",gameIDs[1],"/feed/live?site=en_nhl")))
#first play
temp$liveData$plays$allPlays[[4]]
#probably parse each event type differently
temp$liveData$plays$allPlays[[4]]$result$event
length(temp$liveData$plays$allPlays)

#temp$liveData$plays$allPlays[[363]]
#temp$liveData$plays$allPlays[[364]]
#temp$liveData$plays$allPlays[[365]]
#last one looks like this
# temp$liveData$plays$allPlays[[366]]

temp <- fromJSON(getURL(paste0("https://statsapi.web.nhl.com/api/v1/game/",gameIDs[1],"/feed/live?site=en_nhl")))

a <- lapply(temp$liveData$plays$allPlays, function(x){print(x$`result`$`event`)}) %>% unlist() %>% unique()

temp <- fromJSON(getURL(paste0("https://statsapi.web.nhl.com/api/v1/game/",gameIDs[2],"/feed/live?site=en_nhl")))
b <- lapply(temp$liveData$plays$allPlays, function(x){print(x$`result`$`event`)}) %>% unlist() %>% unique()

temp <- fromJSON(getURL(paste0("https://statsapi.web.nhl.com/api/v1/game/",gameIDs[3],"/feed/live?site=en_nhl")))
c <- lapply(temp$liveData$plays$allPlays, function(x){print(x$`result`$`event`)}) %>% unlist() %>% unique()

unique(c(a,b,c))

eventTypes <- c("Game Scheduled",
                "Period Ready",
                "Period Start",
                "Faceoff",
                "Shot",
                "Stoppage",
                "Blocked Shot",
                "Giveaway",
                "Hit",
                "Missed Shot",
                "Takeaway",
                "Goal",
                "Penalty",
                "Period End",
                "Period Official",
                "Game End")
```

Unique event types I see after checking a few games:
[1] "Game Scheduled"  "Period Ready"   
 [3] "Period Start"    "Faceoff"        
 [5] "Shot"            "Stoppage"       
 [7] "Blocked Shot"    "Giveaway"       
 [9] "Hit"             "Missed Shot"    
[11] "Takeaway"        "Goal"           
[13] "Penalty"         "Period End"     
[15] "Period Official" "Game End" 

Seems like I can just unlist each list element similar to temp$liveData$plays$allPlays[[4]].  Ones without players are missing fields but that should be fine...

```{r}
a <- apply(temp$liveData$plays$allPlays, unlist)
#need to modify names for events with more than 1 player
#otherwise same names and converting to data frame loses them
replaceNames <- function(x) { 
   modify <- c("players.player.id",
                "players.player.fullName",
                "players.player.link", 
                "players.playerType", 
                "players.seasonTotal")
       while("players.playerType" %in% names(x)){
      replace <- which(names(x) %in% modify)
      numReplace <- which(modify %in% names(x))
      if((x["players.playerType"] == "Assist") & ("players.playerType.Assist.1" %in% names(x))){
        temp <- paste(modify[numReplace], x["players.playerType"], "2", sep = ".")
      }  else if(x["players.playerType"] == "Assist") {
        temp <- paste(modify[numReplace], x["players.playerType"], "1", sep = ".")
      } else {
        temp <- paste(modify[numReplace], x["players.playerType"], sep = ".")
      }
      names(x)[replace[numReplace]] <- temp
       }
   return(x)
  }
b <- a %>% lapply(replaceNames) %>%   
  lapply(as.data.frame, stringsAsFactors = FALSE) %>% lapply(t) %>% lapply(as.data.frame, stringsAsFactors = FALSE) %>% bind_rows()

```

Ok try to put it together

```{r}
replaceNames <- function(x) { 
   modify <- c("players.player.id",
                "players.player.fullName",
                "players.player.link", 
                "players.playerType", 
                "players.seasonTotal")
       while("players.playerType" %in% names(x)){
      replace <- which(names(x) %in% modify)
      numReplace <- which(modify %in% names(x))
      if((x["players.playerType"] == "Assist") & ("players.playerType.Assist.1" %in% names(x))){
        temp <- paste(modify[numReplace], x["players.playerType"], "2", sep = ".")
      }  else if(x["players.playerType"] == "Assist") {
        temp <- paste(modify[numReplace], x["players.playerType"], "1", sep = ".")
      } else {
        temp <- paste(modify[numReplace], x["players.playerType"], sep = ".")
      }
      names(x)[replace[numReplace]] <- temp
       }
   return(x)
}

#a <- lapply(temp$liveData$plays$allPlays, unlist)

flattenNHL <- function(x){lapply(x$liveData$plays$allPlays, unlist)}
flattenAddGameID <- function(x){
  flat <- flattenNHL(x)
  lapply(flat, function(y, x){y["gamePk"] <- x$gamePk
  return(y)}, x = x)
  }

parseNHLGame <- function(x){
  x %>% lapply(replaceNames) %>%   
  lapply(as.data.frame, stringsAsFactors = FALSE) %>% lapply(t) %>% lapply(as.data.frame, stringsAsFactors = FALSE) %>% bind_rows()
}


parseNHLPlayByPlay <- function(gameIDs){
  lapply(gameIDs, function(x){fromJSON(getURL(paste0("https://statsapi.web.nhl.com/api/v1/game/", x, "/feed/live?site=en_nhl")))}) %>% lapply(flattenAddGameID) %>% lapply(parseNHLGame) %>% bind_rows() 
  }

playData <- parseNHLPlayByPlay(gameIDs[1:4])

#data <- lapply(gameIDs2, function(x){fromJSON(getURL(paste0("https://statsapi.web.nhl.com/api/v1/game/", x, "/feed/live?site=en_nhl")))}) %>% lapply(flattenAddGameID) %>% lapply(parseNHLPlay) %>% bind_rows() 

```

